---
title: "prop_filled"
output: html_document
date: "`r Sys.Date()`"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(here)
library(tidyverse)
library(magrittr)
```

what are the proportions of ecoregions filled by a parasite given it's host range
break it down by (1) parasite type, (2) transmission mode, (3) host specificity, (4) high vs. low prevalence, (5) high vs. low intensity
```{r}
host_para_mat <- readRDS(here("data", "processed_data", "para_mat.rds"))
ecoreg_host_mat <- readRDS(here("data", "processed_data", "ecoreg_mat.rds"))
ecoreg_para_mat <- readRDS(here("data", "processed_data", "ecoreg_para_mat.rds"))
null_para_df <- readRDS(here("data", "processed_data", "null_para_df.rds"))
gmpd_ecoreg_effort <- readRDS(here("data", "processed_data", "gmpd_ecoreg_effort.rds"))

gmpd <- read_csv(here("data", "GMPD", "GMPD_main.csv"))
```
Add column for parasite type
```{r}
gmpd_type <- gmpd %<>% dplyr::select(ParasiteCorrectedName, ParType) %>% distinct()

par_type_index <- tibble(ParType = c("Virus", "Bacteria", "Protozoa", "Prion", "Fungus", "Helminth", "Arthropod"),
                         ParMacroMicro = c("micro", "micro", "micro", "micro", "micro", "macro", "macro"))

gmpd_type %<>% left_join(., par_type_index)

gmpd_type %<>% rename(., para = ParasiteCorrectedName)
```

```{r}
para_names <- null_para_df %>% dplyr::select(para) %>% distinct()
gmpd_type %<>% filter(para %in% para_names$para)
```

There are some parasites called "not identified to genus" and these are different parasites so are creating issues when joining dataframes
```{r}
gmpd_type %<>% filter(!para == "not identified to genus")
null_para_df %<>% filter(!para == "not identified to genus")

```


```{r}
null_para_df %<>% left_join(., gmpd_type)
```

Add column for transmission mode
```{r}

```


Add column for host specificity


Add column for prevalence


Add column for intensity



```{r}
para_ecoregs <- null_para_df %>% group_by(para) %>% distinct() %>% summarize(tot_ecoreg = n())
```

```{r}
gmpd_para_ecoreg <- ecoreg_para_mat %>% 
  pivot_longer(cols=2:379) %>% 
  filter(value==1) %>%
  group_by(ParasiteCorrectedName) %>%
  distinct() %>%
  summarize(filled_ecoreg = n()) %>%
  rename(para = ParasiteCorrectedName)
  
```






