---
title: "transmission_mode"
output: html_document
date: "`r Sys.Date()`"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(here)
library(tidyverse)
library(magrittr)
```


Look at differences between macroparasite richness and microparasite richness per ecoregion

Macroparasite may be more constrained by ecoregion because abiotic factors should play a larger role in their range because they sometimes have environmental life stages

```{r}
host_para_mat <- readRDS(here("data", "processed_data", "para_mat.rds"))
ecoreg_host_mat <- readRDS(here("data", "processed_data", "ecoreg_mat.rds"))
ecoreg_para_mat <- readRDS(here("data", "processed_data", "ecoreg_para_mat.rds"))
null_para_df <- readRDS(here("data", "processed_data", "null_para_df.rds"))
gmpd_ecoreg_effort <- readRDS(here("data", "processed_data", "gmpd_ecoreg_effort.rds"))

gmpd <- read_csv(here("data", "GMPD", "GMPD_main.csv"))
```

```{r}
gmpd_type <- gmpd %<>% dplyr::select(ParasiteCorrectedName, ParType) %>% distinct()

par_type_index <- tibble(ParType = c("Virus", "Bacteria", "Protozoa", "Prion", "Fungus", "Helminth", "Arthropod"),
                         ParMacroMicro = c("micro", "micro", "micro", "micro", "micro", "macro", "macro"))

gmpd_type %<>% left_join(., par_type_index)

gmpd_type %<>% rename(., para = ParasiteCorrectedName)
```

```{r}
para_names <- null_para_df %>% dplyr::select(para) %>% distinct()
gmpd_type %<>% filter(para %in% para_names$para)
```

There are some parasites called "not identified to genus" and these are different parasites so are creating issues when joining dataframes
```{r}
gmpd_type %<>% filter(!para == "not identified to genus")
null_para_df %<>% filter(!para == "not identified to genus")

```


```{r}
null_para_df %<>% left_join(., gmpd_type)
```

```{r}
null_para_micro <- null_para_df %>% filter(ParMacroMicro == "micro")
null_para_macro <- null_para_df %>% filter(ParMacroMicro == "macro")
count(null_para_micro) + count(null_para_macro) == count(null_para_df) #test
```

```{r, make mat}
make_mat <- function(para_df){
para_df %>% 
  dplyr::select(c(para, name, value)) %>% 
  distinct() %>% 
  pivot_wider(., names_from = name, values_from = value) %>%
  replace(is.na(.), 0)
}
```


```{r}
get_richness_df <- function(para_mat){
  para_rich <- colSums(para_mat[2:ncol(para_mat)])
  
  tmp <- tibble(para = para_rich, ecoreg = names(para_rich))
  
  tmp %>% left_join(., host_rich_df) %>% replace(is.na(.), 0)
}
```

```{r}
host_rich <- colSums(ecoreg_host_mat[2:ncol(ecoreg_host_mat)])
host_rich_df <- tibble(ecoreg = names(host_rich), host = host_rich)
```




# Weight by ecoregion study effort


```{r}
gmpd_ecoreg_effort %<>% rename(., name = ECO_CODE)


gmpd_ecoreg_effort %>% group_by(n) %>% summarize(volume = n()) %>% ggplot(., aes(x=volume)) + geom_histogram(binwidth = 1)
```

do this for micro
```{r}
para_ecoreg_effort <- null_para_micro %>% left_join(., gmpd_ecoreg_effort) 

para_ecoreg_effort %<>% mutate(n = replace_na(n, 0))

ecoreg_count <- para_ecoreg_effort %>% group_by(sci_name, para) %>% summarize(ecoreg_count = n()) %>% ungroup()

para_ecoreg_effort %<>% left_join(., ecoreg_count)

para_ecoreg_effort %<>% mutate(n_times = ifelse(n==0, 1, n)) #turn 0's into 1's for replicate count

effort_weighted_micro <- para_ecoreg_effort %>% uncount(., n_times) #generate row per citation from ecoreg_count
#uncount replicates row per number of times specified
```

do this for macro
```{r}
para_ecoreg_effort <- null_para_macro %>% left_join(., gmpd_ecoreg_effort) 

para_ecoreg_effort %<>% mutate(n = replace_na(n, 0))

ecoreg_count <- para_ecoreg_effort %>% group_by(sci_name, para) %>% summarize(ecoreg_count = n()) %>% ungroup()

para_ecoreg_effort %<>% left_join(., ecoreg_count)

para_ecoreg_effort %<>% mutate(n_times = ifelse(n==0, 1, n)) #turn 0's into 1's for replicate count

effort_weighted_macro <- para_ecoreg_effort %>% uncount(., n_times) #generate row per citation from ecoreg_count
#uncount replicates row per number of times specified
```

effort_weighted is a df where each parasite-host-ecoregion entry is duplicated according to the number of gmpd entries per ecoregion. We duplicate rows so that ecoregions with higher study effort are more likely to yield a presence in the null model because each row is subject to a bernoulli trial so more rows means more opportunities for successul trial (i.e., presence)
```{r}
weighted_prop <- seq(0.01, 0.001, by=-0.0005)
weighted_prop_list <- list()

make_list <- function(effort_weighted){
  for(i in 1:length(weighted_prop)){
    weighted_prop_list[[i]] <- effort_weighted %>% 
    mutate(included = rbinom(value, 1, prob=weighted_prop[i])) %>% 
    filter(included==T) %>%
    distinct()
  }
  names(weighted_prop_list) <- weighted_prop
  return(weighted_prop_list)
}

weighted_prop_micro <- make_list(effort_weighted_micro)
weighted_prop_macro <- make_list(effort_weighted_macro)


```

```{r}
weighted_mat_list_micro <- lapply(weighted_prop_micro, make_mat)
weighted_mat_list_macro <- lapply(weighted_prop_macro, make_mat)

weighted_richness_micro <- lapply(weighted_mat_list_micro, get_richness_df)
weighted_richness_macro <- lapply(weighted_mat_list_macro, get_richness_df)  

make_plots <- function(weighted_richness){
  weighted_plots <- list()
  for(i in 1:length(weighted_richness)){
    weighted_plots[[i]] <- weighted_richness[[i]] %>% ggplot(., aes(x=host, y=para)) + geom_point() + geom_smooth(method = "lm") + labs(title = names(weighted_richness[i])) + xlim(0,65) + ylim(0,270)
  }
return(weighted_plots)
}

weighted_plots_micro <- make_plots(weighted_richness_micro)
weighted_plots_macro <- make_plots(weighted_richness_macro)
```

```{r}
weighted_plots_micro
```

```{r}
weighted_plots_macro
```

```{r}
get_lm <- function(df){
  lm(para ~ host, df)
}
```


```{r}
weighted_lm_micro <- lapply(weighted_richness_micro, get_lm)
weighted_lm_macro <- lapply(weighted_richness_macro, get_lm)
```


```{r}
lapply(weighted_lm_micro, coef)
```

```{r}
lapply(weighted_lm_macro, coef)
```


# Actual GMPD data


```{r}
gmpd_type %<>% rename(ParasiteCorrectedName=para)
ecoreg_para_type <- left_join(ecoreg_para_mat, gmpd_type)

ecoreg_para_mat_micro <- ecoreg_para_type %>% 
  filter(ParMacroMicro=="micro") %>% 
  dplyr::select(-c(ParType, ParMacroMicro))

ecoreg_para_mat_macro <- ecoreg_para_type %>% 
  filter(ParMacroMicro=="macro") %>% 
  dplyr::select(-c(ParType, ParMacroMicro))

actual_para_rich_micro <- get_richness_df(ecoreg_para_mat_micro)
actual_para_rich_macro <- get_richness_df(ecoreg_para_mat_macro)
```




```{r}
actual_para_rich_micro %>% filter(para>0) %>% ggplot(.,aes(x=host, y=para)) + geom_point() + geom_smooth(method="lm") + labs(title = "GMPD microparasites")

summary(lm(para ~ host, data = actual_para_rich_micro))
```

```{r}
actual_para_rich_macro %>% filter(para>0) %>% ggplot(.,aes(x=host, y=para)) + geom_point() + geom_smooth(method="lm") + labs(title = "GMPD macroparasites")

summary(lm(para ~ host, data = actual_para_rich_macro))
```

