---
title: "transmission_mode"
output: html_document
date: "`r Sys.Date()`"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(here)
library(tidyverse)
library(magrittr)
```


Look at differences between complex life cycle and vector borne parasites per ecoregion



```{r}
host_para_mat <- readRDS(here("data", "processed_data", "para_mat.rds"))
ecoreg_host_mat <- readRDS(here("data", "processed_data", "ecoreg_mat.rds"))
ecoreg_para_mat <- readRDS(here("data", "processed_data", "ecoreg_para_mat.rds"))
null_para_df <- readRDS(here("data", "processed_data", "null_para_df.rds"))
gmpd_ecoreg_effort <- readRDS(here("data", "processed_data", "gmpd_ecoreg_effort.rds"))

gmpd <- read_csv(here("data", "GMPD", "GMPD_main.csv"))
gmpd_traits <- read_csv(here("data", "GMPD", "GMPD_parasite_traits.csv"))
```
```{r}
gmpd_traits %<>% rename(para=ParasiteCorrectedName)
```


```{r}
null_para_df %<>% left_join(., gmpd_traits)
```


Make df's for parasites with capacity for direct (close or nonclose) or indirect transmission (no capacity for close or nonclose)

```{r}
null_para_df %<>% mutate(mode = 
                           ifelse(close == TRUE | nonclose == TRUE,
                                  "direct",
                                  "indirect")) %>%
  filter(!is.na(mode)) #remove parasites without mode data

null_para_direct <- null_para_df %>% filter(mode == "direct") #hosts with capacity for close or nonclose transmission
null_para_indirect <- null_para_df %>% filter(mode == "indirect") #hosts without capacity for close or nonclose transmission
count(null_para_direct) + count(null_para_indirect) == count(null_para_df) #test


```

```{r, make mat}
make_mat <- function(para_df){
para_df %>% 
  dplyr::select(c(para, name, value)) %>% 
  distinct() %>% 
  pivot_wider(., names_from = name, values_from = value) %>%
  replace(is.na(.), 0)
}
```


```{r}
get_richness_df <- function(para_mat){
  para_rich <- colSums(para_mat[2:ncol(para_mat)])
  
  tmp <- tibble(para = para_rich, ecoreg = names(para_rich))
  
  tmp %>% left_join(., host_rich_df) %>% replace(is.na(.), 0)
}
```

```{r}
host_rich <- colSums(ecoreg_host_mat[2:ncol(ecoreg_host_mat)])
host_rich_df <- tibble(ecoreg = names(host_rich), host = host_rich)
```




# Weight by ecoregion study effort


```{r}
gmpd_ecoreg_effort %<>% rename(., name = ECO_CODE)


gmpd_ecoreg_effort %>% group_by(n) %>% summarize(volume = n()) %>% ggplot(., aes(x=volume)) + geom_histogram(binwidth = 1)
```

do this for direct
```{r}
para_ecoreg_effort <- null_para_direct %>% left_join(., gmpd_ecoreg_effort) 

para_ecoreg_effort %<>% mutate(n = replace_na(n, 0))

ecoreg_count <- para_ecoreg_effort %>% group_by(sci_name, para) %>% summarize(ecoreg_count = n()) %>% ungroup()

para_ecoreg_effort %<>% left_join(., ecoreg_count)

para_ecoreg_effort %<>% mutate(n_times = ifelse(n==0, 1, n)) #turn 0's into 1's for replicate count

effort_weighted_direct <- para_ecoreg_effort %>% uncount(., n_times) #generate row per citation from ecoreg_count
#uncount replicates row per number of times specified
```

do this for indirect
```{r}
para_ecoreg_effort <- null_para_indirect %>% left_join(., gmpd_ecoreg_effort) 

para_ecoreg_effort %<>% mutate(n = replace_na(n, 0))

ecoreg_count <- para_ecoreg_effort %>% group_by(sci_name, para) %>% summarize(ecoreg_count = n()) %>% ungroup()

para_ecoreg_effort %<>% left_join(., ecoreg_count)

para_ecoreg_effort %<>% mutate(n_times = ifelse(n==0, 1, n)) #turn 0's into 1's for replicate count

effort_weighted_indirect <- para_ecoreg_effort %>% uncount(., n_times) #generate row per citation from ecoreg_count
#uncount replicates row per number of times specified
```

effort_weighted is a df where each parasite-host-ecoregion entry is duplicated according to the number of gmpd entries per ecoregion. We duplicate rows so that ecoregions with higher study effort are more likely to yield a presence in the null model because each row is subject to a bernoulli trial so more rows means more opportunities for successul trial (i.e., presence)
```{r}
weighted_prop <- seq(0.01, 0.001, by=-0.0005)
weighted_prop_list <- list()

make_list <- function(effort_weighted){
  for(i in 1:length(weighted_prop)){
    weighted_prop_list[[i]] <- effort_weighted %>% 
    mutate(included = rbinom(value, 1, prob=weighted_prop[i])) %>% 
    filter(included==T) %>%
    distinct()
  }
  names(weighted_prop_list) <- weighted_prop
  return(weighted_prop_list)
}

weighted_prop_direct <- make_list(effort_weighted_direct)
weighted_prop_indirect <- make_list(effort_weighted_indirect)


```

```{r}
weighted_mat_list_direct <- lapply(weighted_prop_direct, make_mat)
weighted_mat_list_indirect <- lapply(weighted_prop_indirect, make_mat)

weighted_richness_direct <- lapply(weighted_mat_list_direct, get_richness_df)
weighted_richness_indirect <- lapply(weighted_mat_list_indirect, get_richness_df)  

make_plots <- function(weighted_richness){
  weighted_plots <- list()
  for(i in 1:length(weighted_richness)){
    weighted_plots[[i]] <- weighted_richness[[i]] %>% ggplot(., aes(x=host, y=para)) + geom_point() + geom_smooth(method = "lm") + labs(title = names(weighted_richness[i])) + xlim(0,65) + ylim(0,270)
  }
return(weighted_plots)
}

weighted_plots_direct <- make_plots(weighted_richness_direct)
weighted_plots_indirect <- make_plots(weighted_richness_indirect)
```

```{r}
weighted_plots_direct
```

```{r}
weighted_plots_indirect
```

```{r}
get_lm <- function(df){
  lm(para ~ host, df)
}
```


```{r}
weighted_lm_direct <- lapply(weighted_richness_direct, get_lm)
weighted_lm_indirect <- lapply(weighted_richness_indirect, get_lm)
```


```{r}
lapply(weighted_lm_direct, coef)
```

```{r}
lapply(weighted_lm_indirect, coef)
```


# Actual GMPD data


```{r}
gmpd_traits %<>% rename(ParasiteCorrectedName = para)
ecoreg_para_mode <- left_join(ecoreg_para_mat, gmpd_traits)

ecoreg_para_mode %<>% mutate(mode = 
                           ifelse(close == TRUE | nonclose == TRUE,
                                  "direct",
                                  "indirect")) %>%
  filter(!is.na(mode)) #remove parasites without mode data


ecoreg_para_mat_direct <- ecoreg_para_mode %>% 
  filter(mode=="direct") %>% 
  dplyr::select(-c(close:mode))

ecoreg_para_mat_indirect <- ecoreg_para_mode %>% 
  filter(mode=="indirect") %>% 
  dplyr::select(-c(close:mode))

actual_para_rich_direct <- get_richness_df(ecoreg_para_mat_direct)
actual_para_rich_indirect <- get_richness_df(ecoreg_para_mat_indirect)
```




```{r}
actual_para_rich_direct %>% filter(para>0) %>% ggplot(.,aes(x=host, y=para)) + geom_point() + geom_smooth(method="lm") + labs(title = "GMPD direct parasites")

summary(lm(para ~ host, data = actual_para_rich_direct))
```

```{r}
actual_para_rich_indirect %>% filter(para>0) %>% ggplot(.,aes(x=host, y=para)) + geom_point() + geom_smooth(method="lm") + labs(title = "GMPD indirect parasites")

summary(lm(para ~ host, data = actual_para_rich_indirect))
```

